---
title: "Index.Rmd"
author: "Jonathan Schuster"
date: "08/10/2020"
output: html_document
---
 <!-- This initial chunk is loading all of the packages I'll be utilizing for my COVID report. Example, I'll be using the package "choroplethr" to set up choropleth maps as a form of graph. --> 

```{r setup, include=FALSE, echo = FALSE}
library(tidyverse)
library(stringr)
library(choroplethr)
library(choroplethrMaps)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

 <!-- This chunk is uploading all the databases to my Rmd file. Files are from the CDC. I elected to use two different datasets from the CDC and you'll see below the ways I've elected to manipulate the datasets chosen.   --> 

```{r importing, message = FALSE, echo = FALSE}
data(state.regions)
myData <- read_csv("raw_data/US_MAP_DATA.csv", skip = 2)
myData2 <- read_csv("raw_data/Provisional_COVID-19_Death_Counts_by_Place_of_Death_and_State.csv")

```

 <!-- In this chunk I am manipulating a data set into comparison states and making a choropleth representation. I'm renaming "Total Death" to "Deaths" for simplicity in writing my code just like with my Midwest states. I've filtered my data set into my 7 similar pop states and changed the case to lower case with the str_to_lower function on jurisdiction. Lastly, I'm saving this csv into my output folder now that I've manipulated the raw data. --> 

```{r tidyup2, message=FALSE, echo=FALSE, warning=FALSE}

deathData2 <- myData %>% 
        rename(Deaths = "Total Death") %>%
        filter(fips < 57, abbr != "US" & jurisdiction %in% c("New Jersey", "Georgia", "Maryland", "Colorado", "Pennsylvania", "Tennessee", "Louisiana")) %>%
        summarise(region = str_to_lower(jurisdiction), value = Deaths)

write.csv(deathData2,"C:/Users/Jonny/Documents/GitHub/Assignment-10/output/deaths_by_state_similar_pop.csv", row.names = FALSE)

```

 <!-- In this chunk I'm taking my manipulated data and making my choropleth map for my selected comparison states. I've given the graph a nice title and labeled my legend "Deaths" to give viewers a way to decipher my image. --> 

```{r choropleth2, message=FALSE, echo=FALSE, warning=FALSE, out.width = "500px", fig.align='center', fig.cap="**Choropleth 2:** Total Deaths from Comparison States"}

state_choropleth(deathData2, title = "Total Deaths due to COVID", legend = "Deaths")

```

**Description:**

This dataset comes from the CDC and the direct url is https://www.cdc.gov/covid-data-tracker/#cases. Same manipulations occur in this choropleth graph as the one presented in assignment 09. The only differences is I used the similar pop states rather than the midwest coalition states. This gives the viewer a nicer graph to look at just like the other choropleth. I'll be using these choropleths to show the viewer the total death counts to visualize how coalition states saw less deaths overall than noncoalition states.